\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amsthm, amssymb, mathpazo, isomath, mathtools}
\usepackage{subcaption,graphicx,pgfplots}
\usepackage{fullpage}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{algorithm, algorithmic}
\usepackage{mathtools}
\usepackage{todonotes}

\newcommand{\example}[1]{\todo[inline,color=green!30!white]{\textbf{Example:} #1}}

\title{Notes for Advanced Topics in Mathematical Engineering}
%\author{Nicol\'as A Barnafi\thanks{Instituto de Ingeniería Biológica y Médica, Pontificia Universidad Católica de Chile, Chile}, Axel Osses\thanks{Departamento de Ingeniería Matemática, Universidad de Chile, Chile}}
\author{Nicol\'as A Barnafi}
%\date{}

\renewcommand{\vec}{\vectorsym}
\newcommand{\mat}{\matrixsym}
\newcommand{\ten}{\tensorsym}
\DeclareMathOperator{\grad}{\nabla}
\DeclareMathOperator{\dive}{\text{div}}
\DeclareMathOperator{\curl}{\text{curl}}
\newtheorem{remark}{Remark}
\newtheorem{definition}{Definition}
\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathcal{D}}

\newcommand{\tin}{\text{in}}
\newcommand{\ton}{\text{on}}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\usepackage{listings}
\usepackage{xcolor}
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
  backgroundcolor=\color{backcolour}, commentstyle=\color{codegreen},
  keywordstyle=\color{magenta},
  numberstyle=\tiny\color{codegray},
  stringstyle=\color{codepurple},
  basicstyle=\ttfamily\footnotesize,
  breakatwhitespace=false,         
%   breaklines=false,                     
  captionpos=b,                    
  keepspaces=true,                 
  numbers=none,                    
  numbersep=5pt,                  
  showspaces=false,                
  showstringspaces=false,
  showtabs=false,                  
  tabsize=2,
%   frameround=tttn,
  framerule=1.5pt,
  rulecolor=\color{red!60!black}
}
\lstset{style=mystyle}

\begin{document}

\maketitle

\section*{Context}

These notes exist as backup material for a course on some deeper topics in Math Eng at Pontificia Universidad Católica de Chile, the 2nd semester of 2024. The idea is to provide mathematical tools for students that give them the ability to assess the difficulty of mathematical problems, mainly within the world of Partial Differential Equations (PDEs). The target is ultimately to implement these models, so that all tools are oriented towards having solid foundations that allows one to trust a computational model. Informally speaking, the main mathematical concepts to haunt us during all these notes are: 
    \begin{itemize}
        \item Existence and uniqueness: It is a natural baseline in the mathematician's world to try to solve only problems that \emph{have} a solution. Otherwise, things might be as pointless as developing an iterative method for finding real numbers such that $x^2 = -1$. Uniqueness is a further luxury, but sometimes two different methods give two different solutions, and having only those things at hand can make it difficult to distinguish whether that is a bug or a feature of the model. There exist some root-isolation methods that allow to find solutions of a problem that are \emph{different} from a given one. This is out of the scope of this course. 
        \item Stability: The intuitive idea behind this is that small perturbations in the data give rise to small changes in the solution. This typically looks like 
            $$ \| u\|_X \leq \| f\|_{X'}, $$
        where $u$ is the solution of a problem that depends on $f$, and $X$ is some functional (hopefully Hilbert) space. More rigorously, this means that the solution map $f \mapsto u(f)$ is bounded, or continuous in the linear case. Stability also sometimes refers to time dynamics and the fact that a discrete solution stays \emph{within a certain distance} of the real solution throughout a simulation. In the continuous setting, it might also mean that there are no finite-time singularities. In general, stability is not a well defined term, but still a widely understood one to anyone who has struggled to get a code to run correctly, and a highly desired property. 
    \end{itemize}
All other properties (or at least most of them anyway) are ways to guarantee that a problem enjoys one of these nice properties. There are ways to handle problems that do not have those properties, but they are almost always extremely problem dependent, and the person studying such problems should dive deep into the sectorial knowledge to see how certain communities deal with such issues. This is an aspect that mathematically oriented people almost always disregard, which has some severe mathematical (and social) consequences. In fact, some extremely classical models in engineering are still far from understood mathematically, such as the Navier-Stokes equations. This has not prevented the CFD community from solving these models with extreme efficiency, and from further leveraging them for industrial applications which, unsurprisingly, work fantastically. Discovering the amazing ways in which mathematically obvlivious communities solve mathematically hard problems is, and will probably be for very long, a beautiful opportunity for collaboration.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis preliminaries}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section we will review some important properties of functional spaces and operators. These things should be deemed as 'review' material. Intrinsically new things will start appearing in Secion~\ref{section:beyond-ellipticity}. Most, if not all, results will be coming from the amazing book \emph{Linear and nonlinear functional analysis} by PG Ciarlet.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Functional spaces}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Banach and Hilbert spaces} Throughout the entire manuscript, we will rely on Banach spaces, Hilbert spaces, and their duals. Despite the existence of a flexible theory of Banach space formulations, we will mostly rely on Hilbert spaces because of their many nice properties. For now, let's simply review some relevant properties: 
    \begin{itemize}
        \item Banach spaces are complete metric spaces. For a given Banach space $X$, its (topological) dual is the space $X'$ of functions $X\mapsto \R$. The action of an element in the dual space is sometimes denoted as $\langle T, x\rangle_{X'\times X}$, so as to resemble the notation of an inner product. In general, one can identify a part of the bidual space $X''$ through the evaluation operator $T_f:X'\mapsto \R$ in $X''$ defined as $T_f(L) = L(f)$. This immersion is not surjective. 
        \item Continuous linear operators acting on Banach spaces have an induced norm: If $T: X\mapsto Y$, then 
            $$ \| T\| =  \sup_{x\in X}\frac{|Tx|_Y}{|x|_X}. $$
        Some people write this space as $L(X,Y)$. 
        \item Hilbert spaces are Banach spaces with respect to the distance induced by a dot (inner) product, i.e. a bilinear form $\langle\cdot, \cdot \rangle: X\times X\mapsto \R $ such that: 
            \begin{itemize}
                \item It is symmetric: $\langle x,y\rangle = \langle y, x\rangle$
                \item It is linear in its first argument: $\langle \alpha x_1 + \beta x_2, y\rangle=\alpha\langle x_1, y\rangle + \beta\langle x_2, y\rangle$
                \item It is positive definite: $\langle x,x\rangle \geq 0$, where it is 0 only if $x=0$.
            \end{itemize}
        \item The inner product yields the fantastic Riesz map, which is actually an isometry. This is given as follows: Consider a Hilbert space $H$ with inner product $\langle\cdot, \cdot\rangle_H$, then a Riesz map is an operator $R_H: H\mapsto H'$ such that for any $x,y$ in $H$ it holds that $\langle R_H(x), y\rangle_{H'\times H} = \langle x, y\rangle_H$. Notably, $\|R_H(x)\|_{H'} = \| x \|_H$. 
        \item Inner products are mostly used as projections. This means that, in the same way that we can orthogonalize a vector $x$ with respect to $y$, we can also do this in the Hilbert space setting analogously as 
            $$ x_\perp \coloneqq x - \langle x, y\rangle_H y. $$
        It can be quickly verified that the function $x_\perp$ is indeed perpendicular to $y$ in the sense that $\langle x_\perp, y\rangle_H=0$. 
    \end{itemize}
One fundamental aspect of Hilbert spaces is that they provide some intuitive properties related to projections, which we recall through the following results: 
\begin{theorem}{Best approximation}
    Set $U\subset H$ a closed subspace of a Hilbert space $H$ and set $f$ in $H$. Then there exists a unique $g$ in $U$ such that
        $$ \|f - g \|_H = \inf_{u\in U} \| f - u\|_H. $$
\end{theorem}
Using this, we can uniquely definte the orthogonal complement of a set $U$: 
    $$ U^\perp \coloneqq \{v\in H: (v, u)_H = 0 \quad\forall u\in U\}. $$
\begin{theorem}
    Set $U$ a closed subspace of a Hilbert space $H$. Then, if $f$ is in $H$, there exists a unique pair $(u,v)$ in $U\times U^\perp$ such that 
        $$f = u + v.$$
\end{theorem}
Some common and/or simple examples: Helmholtz decomposition, zero average functions, zero trace tensors, symmetric tensors. Note that the orthogonal complement is defined with respect to a \emph{given} inner product.

\example{
  Assume we want to orthogonalize with respect to $U=\R$. Then, we have that there is a constant $c$ such that for $f$ in a Hilbert space $H$ we can write
    $$ f = h + c, $$
  with $h\perp U$. Noting that a function $x$ satisfies $x\perp U$ iff $(x,1)_H = 0$, then we can use the previous expression to obtain 
    $$ (f,1)_H = (c,1) = c(1,1)_H, $$ 
  which gives
    $$ c = \frac{(f,1)_H}{(1,1)_H}. $$
}

The most important spaces for us will be the Lebesgue spaces $L^p(\Omega;\R^d)$ given by measurable functions $f:\Omega \mapsto \R^d$ such that
    $$ \int_\Omega |f|_{\R^d}^p\,dx < \infty. $$
It will be important to know that if $|\Omega|<\infty$, then these spaces form an ordered inclusion: 
    $$ L^\infty(\Omega) \subset L^p(\Omega) \subset ... \subset L^1(\Omega). $$
A simple way to remember this is to split a function as $f = I_{|f|\leq 1}f + I_{|f|\geq 1}f$ and note that $|x|^p < |x|^{p+\epsilon}$ for $\epsilon > 0$. 

\paragraph{Distributions and derivatives} To formulate differential equations in Banach/Hilbert spaces, it will be important to be able to define derivatives in such spaces. This is done through the language of distributions, invented (discovered) by L Schwartz. For this, we require the notion of 'test functions', i.e. functions on which we can discharge derivatives of abstract objects through integration by parts. Consider then a function $f$ in $C_0^\infty(\R^d)$, the space of infinitely differentiable scalar functions with compact support in $\R^d$, then a distribution is simply an element $T$ in the dual space $(C_0^\infty(\R^d))'$, whose action can be written as $\langle T, f\rangle_{(C_0^\infty)'\times C_0^\infty}$, or sometimes simply as $\langle T, f\rangle$, if it is clear by context. The idea is to generalize the notion of action through integration, so that for sufficiently smooth functions $f$, their induced distribution is $Tf$ given by
    $$ \langle Tf, g \rangle = \int_{\R^d} fg\,dx. $$
This integral approach, when thinking about integration by parts formulas, allows us to define distribution derivatives as
    $$ \langle \partial_i T, f\rangle \coloneqq -\langle T, \partial_i f\rangle, $$
as given by integration by parts. This is known as a \emph{weak derivative}. Arbitrary order differential operators can be defined analogously, most importantly $\grad, \dive, \curl$, given by 
    \begin{align*}
        \langle \dive T, f\rangle &\coloneqq -\langle T, \grad f\rangle \\
        \langle \curl T, f\rangle &\coloneqq \langle T, \curl f\rangle.
    \end{align*}
\example{All classical (or strong) derivatives coincide with the weak derivatives, as seen from the integration by parts formulas. Also, consider the Dirac delta distribution given by
    $$ \langle \delta_x, f\rangle = f(x), $$
sometimes written as $\delta_x(f)$, or also simply as $\int_\Omega \delta_x f\,dx$ (with a \emph{not too mild} abuse of notation).
Then, its derivative is given by 
    $$ \langle \delta', f \rangle = -\langle \delta_x, f'\rangle = - f'(x).$$
}
Naturally, weak derivatives and the common ones coincide under differentiability assumptions. The notion of weak derivatives allows us to define differentiable Hilbert spaces, given by 
    $$ W^{1,p}(\Omega) \coloneqq \{ f\in L^p(\Omega): \grad f \in L^p(\Omega)\}. $$
These are Banach spaces with norm
    $$ \| x \|_{W^{1,p}(\Omega)} \coloneqq \| x \|_{L^p(\Omega)} + \| \grad x \|_{L^p(\Omega)}. $$
This is the graph norm of the $\nabla$ operator, and it can be further seen as the $\ell^1$ norm of the two-dimensional vector $(\|x\|_{L^p(\Omega)}, \|\grad x\|_{L^p(\Omega)})$, and thus all vector norms for such a vector induce equivalent norms for Sobolev spaces. It is very common to use the following notations:   
    \begin{itemize}
        \item $H^1(\Omega) = W^{1,2}(\Omega)$.
        \item $\| x \|_{L^2(\Omega)} = \| x \|_{0,\Omega}$, or even simply $ \| x\|_0$, depending on the laziness of the person writing. 
        \item $\| x \|_{H^1(\Omega)} = \|x\|_{1,\Omega}$.
    \end{itemize}
The space $H^1(\Omega)$ is very important, as it is a Hilbert space with inner product
    $$ \langle x,y\rangle_{H^1(\Omega)} \coloneqq \langle x,y\rangle_{L^2(\Omega)} + \langle \grad x, \grad y\rangle_{L^2(\Omega)}. $$
Analogously, we can define the spaces
    $$ H(\dive; \Omega) = \left\{f\in L^2(\Omega): \dive f \in L^2(\Omega)\right\} $$
and    
    $$ H(\curl; \Omega) = \left\{f\in L^2(\Omega): \curl f \in L^2(\Omega)\right\}$$
Their application depends on the context, so we only keep here their definition. They are also Hilbert spaces, with the inner product defined as the one in $H^1$ but with the corresponding differential operators. Note that $H^1$ functions belong to both $H(\dive)$ and $H(\curl)$, but inclusions among them are not clear. We conclude this section with the celebrated Sobolev embedding results: 

\begin{theorem}[Sobolev embeddings (continuous)]
    Consider $\Omega$ a bounded Lipschitz domain in $\R^d$, set $m,j$ two non-negative integers, and $p$ in $[1,\infty]$. Then the following embeddings hold: 
    \begin{enumerate}
        \item If $mp < d$ then for $p \leq q \leq \frac{dp}{d-mp}$:
            $$ W^{j+m, p}(\Omega) \hookrightarrow W^{j,q}(\Omega). $$
        \item If $mp = d$, then for $p \leq q < \infty$:
            $$ W^{j+m, p}(\Omega) \hookrightarrow W^{j,q}(\Omega). $$
        \item If $mp > d \geq (m-1)p$, then 
            $$ W^{j+m,p}(\Omega) \hookrightarrow C^j(\bar\Omega). $$
    \end{enumerate}
\end{theorem}

\begin{theorem}[Sobolev embeddings (compact)]
Consider $\Omega$ a bounded Lipschitz domain in $\R^d$, $m\geq 1$ integer, $j\geq 0$ integer and let $p$ in $[1,\infty)$. Then the following embeddings are compact:
    \begin{enumerate}
        \item If $mp \leq d$  then for $q$ in $[1, \frac{dp}{d - mp})$:
            $$ W^{j+m, p}(\Omega) \hookrightarrow W^{j,q}(\Omega). $$
        \item If $mp > d$, then 
            $$ W^{j+m, p}(\Omega \hookrightarrow C^j(\bar\Omega). $$
    \end{enumerate}
These inclusions hold also if the arrival domain is an arbitrary subdomain of $\Omega$. 
\end{theorem}
These theorems hold in greater generality, and are typically used together with the weak-compactness of the unit ball to show the existence of certain strongly convergent subsequence. We will see examples of this further ahead. One fundamental consequence is that $H^1$ is compactly embedded in $L^2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Traces}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Traces or trace operators are the ones that restrict a function in $\R^d$ to some set in $\R^{d-1}$, most commonly the boundary of a domain. They are fundamental to adequately define boundary conditions. For the presentation of this section, we follow \cite{gatica2014simple} and \cite{monk2003finite}. Some details about Sobolev spaces are drawn from \cite{adams2003sobolev}. The fundamental difficulty of defining trace operators is that the domain where the boundary condition is defined has measure 0 in the measure of the starting domain, so some regularity of the function is required to guarantee that this operation makes sense. We will not enter the details of how a Lipschitz boundary is defined, see \cite{monk2003finite} for further details.

There are several definitions and constructions here that are needed for everything to make sense. We will follow them in a reasonable order, but this might be a very personal vision, so please read other formulations to have a more well-rounded vision. We will denote with $C_0^\infty(X)$ the space of functions with compact support in $X$, and also with $\D(\bar\Omega)$ the functions in $C_0^\infty(\R^d)$ with support in an open set $U$ such that $\bar\Omega\subset U$. This belongs to a wider set known as the Schwartz class of functions. If the set $X$ is open, we may denote $\D(X)$ as $C_0^\infty(X)$ with a bit of an abuse of notation.

\begin{itemize}
    \item Classic densities: $\D(\bar\Omega)$ is dense in $L^p(\Omega)$ if $\Omega$ is bounded and Lipschitz.
    \item $C^\infty(\bar\Omega)$ is dense in $W^{s,p}(\Omega)$ for $s$ a positive integer and $p\in [1,\infty)$.
    \item For $s$ a positive integer and $p\in (1,\infty)$, we have that there is a continuous linear extension $\Pi: W^{s,p}(\Omega) \to W^{s,p}(\R^d)$ such that $\Pi u|_\Omega = u$ for all $u$ in $W^{s,p}(\Omega)$.
\end{itemize}

Some technicalities arise when $\Omega$ is unbounded. For the sake of this course, all domains are bounded and Lipschitz unless otherwise stated. The following theorem allows us to extend the trace operator, which we initially define as $\gamma_0: \D(\bar\Omega) \to C^\infty(\partial\Omega)$, given by
    $$ \gamma_0 u = u|_{\partial\Omega}.$$
An important property that we will use many times is the \emph{trace inequality}, which states that there exists $C$ positive such that 
    $$ \| \gamma_0 f \|_{0,\partial\Omega} \leq C \| f \|_{1,\Omega} \qquad\forall f \in \mathcal D(\bar\Omega). $$

\begin{theorem}[Trace theorem]\label{thm:trace-theorem}
    Set $\Omega$ a bounded and Lipscthiz domain. Then, considering $1/p < s \leq 1$, there exists a continuous extension of $\gamma_0$ given by $\gamma_0: W^{s,p}(\Omega) \to W^{\frac{s-1}{p}, p}(\partial\Omega)$
    \begin{proof}
        We refer to \cite{adams2003sobolev} for a complete proof. We will simply see how to extend $\gamma_0$ from smooth functions to a linear bounded operator from $H^1(\Omega)$ to $L^2(\partial\Omega)$. This result requires the density of $\mathcal D$ in $H^1$ and the trace inequality. Consider thus a Cauchy sequence $\{\varphi_i\}_i$ in $\mathcal D(\bar\Omega)$ converging to some $v$ in $H^1(\Omega)$. Using linearity and continuity, we get that for some pair of indexes $i,j$: 
        $$ \| \gamma_0 \varphi_i - \gamma_0 \varphi_j \|_{0,\partial\Omega} = \| \gamma_0 (\varphi_i - \varphi_j) \|_{0,\partial\Omega} \leq C \| \varphi_i - \varphi_j \|_{1,\Omega}. $$
        This states that $\{ \gamma_0 \varphi_i \}_i $ is a Cauchy sequence in $L^2$, and thus has a limit $\xi$ in $L^2(\partial\Omega)$. Before setting $\xi$ as our extension, we need to check it is independent of the chosen sequence. This can be simply done by choosing another sequence such that $\{\tilde \varphi_i\}_i$ converges also to $v$. Then, it holds that 
        $$ \| \gamma_0 \tilde \varphi_i - \xi \| \leq \| \gamma_0(\tilde\varphi_i - \varphi_i) + \gamma_0 \varphi_i - \xi \| \leq \| \gamma_0(\tilde\varphi_i - \varphi_i) \| + \| \gamma\varphi_i - \xi \|. $$
        The second term goes to zero as we showed previously. For the first one, we use the trace inquality again to obtain 
        $$ \| \gamma_0(\varphi_i - \tilde\varphi_i) \| \leq C \|\varphi_i - \tilde \varphi_i \|, $$
        which concludes the proof. 
    \end{proof}
\end{theorem}

This trace is sometimes referred to as the \emph{Dirichlet} trace, as it is used to define Dirichlet boundary conditions. We can now define the Sobolev spaces "with boundary conditions", i.e. 
    $$ W_0^{1,p} = \{ u \in L^p(\Omega): \grad u \in [L^p(\Omega)]^d \text{ and } \gamma_0 u = 0 \}. $$
Here, we used the standard notation $[X(\Omega)]^d \coloneqq X(\Omega)\times \hdots \times X(\Omega)$. We thus get the extensively used spaces: 
    $$
    \begin{aligned}
        H_0^1(\Omega) &= W_0^{1,2}(\Omega) \\
        H^{-1}(\Omega) &= [H_0^1(\Omega)]' \\
        H^{1/2}(\partial\Omega) &= W^{1/2,2}(\partial\Omega) \\
        H^{-1/2}(\partial\Omega) &= [H^{1/2}(\partial\Omega)]',
    \end{aligned}
    $$
where the space $H^{1/2}(\partial\Omega)$ is the trace space associated to $H^1(\Omega)$, and the kernel of $\gamma_0$ is given by $H_0^1(\Omega)$. 

\paragraph{The trace spaces} This space has some nice properties, which we detail now. Its norm\footnote{Sobolev spaces of fractional order are a best on their own. The rigorous definition can be given using either Fourier transforms or by using the Slobodeckij seminorm. Both are very cumbersome and seldom used.} is given by 

    $$ \| u \|_{1/2,\partial\Omega} \coloneqq \inf\{\|U\|_{1,\Omega}: U \in H^1(\Omega) \text{ and } u = \gamma_0 U\}, $$
which naturally yields the following continuity estimate for the trace operator: 
    $$ \| \gamma_0 U\|_{1/2,\partial\Omega} \leq \| U \|_{1,\Omega} . $$
The trace space $H^{1/2}$ can be seen as a quotient space derived from $H^1$, so it is also a Hilbert space. A natural question is what the inner product looks like. To do that, we consider for a given $u$ in $H^{1/2}(\partial\Omega)$, an element that yields the norm, i.e. $U$ in $H^1(\Omega)$ such that $\gamma_0 U = u$ and $\| u \|_{1/2,\partial\Omega} = \| U \|_{1,\Omega}$. In such a case, we can consider the following inner product: 
    $$ (v_1, v_2)_{1/2,\partial\Omega} \coloneqq (V_1, V_2)_{1,\Omega}, $$
where $V_i$ are the extension functions. Finally, it will be useful to know that $H_0^1(\Omega)$ (and indeed also $W_0^{s,p}$) can be defined as a closure in terms of the $H^1$ norm: 
    $$ H_0^1(\Omega) \coloneqq \overline{C_0^\infty(\Omega)}^{\|\cdot \|_{1,\Omega}}. $$

\paragraph{Note on integration by parts formulas} These formulas will be important to define the normal and tangential traces. All formulas stem from the divergence theorem: 

\begin{theorem}[Divergence Theorem]
    Consider a bounded Lipschitz domain $\Omega$ in $\R^{d=2,3}$ and consider a vector field $\vec F:\R^d \to \R^d$ in $[C^1(\bar\Omega)]^d$. Then it holds that
        $$ \int_\Omega \dive \vec F\,dx = \int_{\partial\Omega}\vec F\cdot \vec n\,ds,$$
    where $\vec n$ is the outwards normal vector, $dx$ is the volume measure and $ds$ is the surface measure.
\end{theorem}

The relevant formulas are the following: 
    \begin{itemize}
        \item If $\xi$ in $C^1(\bar\Omega)$ and $\vec u$ in $[C^1(\bar\Omega)]^d$:
            $$ \int_\Omega (\dive \vec u) \xi\,dx = -\int_\Omega\vec u\cdot \grad \xi\,dx + \int_{\partial\Omega} \vec u \cdot \vec n \xi\,ds.$$
        \item (Green's [first] identity)\footnote{See \cite{monk2003finite} for the second one. It is useful to derive Boundary Element (BEM) methods.} If $\xi$ in $C^1(\bar\Omega)$ and $p$ in $C^2(\bar\Omega)$:
            $$ -\int_\Omega (\Delta p) \xi\,dx = \int_\Omega\grad p\cdot \grad \xi\,dx - \int_{\partial\Omega}\left(\grad p\cdot \vec n\right)\xi\,ds.$$
        \item Consider $\vec u,\vec \phi$ in $[C^1(\bar\Omega)]^d$: 
            $$ \int_\Omega (\curl \vec u) \cdot \vec\phi\,dx = \int_\Omega \vec u \cdot (\curl \vec \phi)\,dx  + \int_{\partial\Omega}(\vec n\times \vec u)\cdot \vec\phi\,ds.$$
    \end{itemize}

\paragraph{$H(\dive)$ and the normal trace} In this space, we have a first simple density result: 

\begin{theorem} Consider a bounded and Lipschitz domain $\Omega$ in $\R^d$, then $H(\dive;\Omega)$ is the closure of $[C(\bar\Omega)]^d$ in the $H(\dive)$ norm.
    \begin{proof}
        Sketch: The main idea is to show that the orthogonal complement of $[C(\bar\Omega]^d$ in $H(\dive;\Omega)$ is the trivial one. Then, one uses the orthogonality condition 
            $$ (\vec u, \vec \phi) + (\dive \vec u, \dive \vec \phi) = 0$$
        for all $\vec\phi$ in $[C(\Omega)]^d$ implies that $\grad \dive\vec u = \vec u$ is in $L^2$, and thus $\dive \vec u$ is in $H^1$. An adequate extension to all $\R^d$ and a density argument concludes the proof. For details, see \cite[Thm 3.22]{monk2003finite}. 
    \end{proof}
\end{theorem}
The normal trace is simply given for a smooth function as 
    $$ \gamma_N \vec v = \vec v|_{\partial\Omega} \cdot n, $$
where $\vec n $ is the outwards normal vector. This can be extended up to functions in $H(\dive)$ as stated in the following theorem. 

\begin{theorem}[Normal trace]
Consider a bounded Lipschitz domain $\Omega$ in $\R^d$ with outwards unit normal $\vec n$. Then, the mapping $\gamma_N$ can be extended to a continuous linear map $\gamma_N: H(\dive; \Omega) \to H^{-1/2}(\partial\Omega)$, and the following integration by parts formula holds: 
        $$ \langle \gamma_N \vec v, \phi \rangle_{-1/2, 1/2} = (\vec v, \grad \phi) + (\dive \vec v, \phi) \qquad \forall \vec v\in H(\dive;\Omega), \phi \in H^1(\Omega). $$
    \begin{proof}
        First, we note that the integration by parts formula, which holds for $C^\infty$ functions initially, can be extended by density to functions $\phi$ in $H^1(\Omega)$:
        $$ \langle \vec v\cdot \vec n, \phi \rangle = (\vec v, \grad \phi) + (\dive \vec v, \phi) \qquad \forall \vec v\in H(\dive;\Omega), \phi \in H^1(\Omega), \qquad \forall \vec v\in H(\dive;\Omega), \phi \in H^1(\Omega). $$
        Cauchy-Schwartz further yields
            $$ |\langle \vec v\cdot \vec n, \phi\rangle | \leq \| \vec v \|_{\dive} \| \phi \|_1. $$
        In particular, using the surjectivity of the Dirichlet trace, we get that for all $\mu$ in $H^{1/2}(\partial\Omega)$ it also holds that
            $$ |\langle \vec v\cdot \vec n, \mu\rangle | \leq \| \vec v \|_{\dive} \| \mu \|_{1/2}, $$
        which naturally yields
            $$ \| \vec v \cdot \vec n\|_{-1/2} \leq \| \vec v\|_{\dive}. $$
        This all implies that $\gamma_N$ is a bounded linear map from $[C(\bar\Omega)]^d$ to $H^{-1/2}(\partial\Omega)$, meaning that it can be extended by density to be defined in $H(\dive;\Omega)$ as well. We are only missing the surjectivity, for which we consider a generic function $\eta$ in $H^{-1/2}(\partial\Omega)$, and define the following problem: Find $\phi$ in $H^1(\Omega)$ such that
        $$ (\grad \phi, \grad \psi) + (\phi, \psi) = \langle \eta, \gamma_0 \psi \rangle_{-1/2,1/2} \qquad\forall \psi \in H^1(\Omega).$$
        This in particular implies that 
        $$ (\grad \phi, \grad \psi_0) + (\phi, \psi_0) = 0 \qquad\forall \psi_0 \in H_0^1(\Omega),$$
        so that, as in the previous proof, $-\dive \grad \phi = \phi$ distributionally and thus $\vec v \coloneqq \grad \phi$ is the $H(\dive;\Omega)$ function we were looking for.
    \end{proof}
\end{theorem}
Finally, we will use the space of functions with null normal trace, so that we initially define
    $$ H_0(\dive, \Omega) \coloneqq \overline{[C_0^\infty(\Omega)]^d}^{\|\cdot \|_{\dive}}, $$
i.e. the closure in the $\dive$ norm. The following result makes all the trouble worth it. 
\begin{theorem}
    Consider a bounded Lipschitz domain $\Omega$ in $\R^d$. Then: 
        $$ H_0(\dive;\Omega) = \{\vec v\in H(\dive;\Omega): \gamma_N \vec v = 0 \}. $$
    \begin{proof}
        The proof is done through the orthogonal complement. All techniques have been already shown here, so we simply refer the interested reader to \cite[Thm 3.25]{monk2003finite}.
    \end{proof}
\end{theorem}

\paragraph{$H(\curl)$ and the tangential trace} In the proofs involving $H(\dive)$, we heavily used the fact that the $\grad$ and $\dive$ are the transpose of one another. This is not the case for the $\curl$ operator, so the proofs for this case are notoriously more complicated. Instead, we will be happy with simply stating the related results:
    \begin{itemize}
        \item $$ H_0(\curl, \Omega) \coloneqq \overline{[C_0^\infty(\Omega)]^d}^{\|\cdot\|_{\curl}}. $$
        \begin{theorem}
            For a bounded Lipschitz domain $\Omega$, it holds that $[C(\bar\Omega)]^3$ is dense in $H(\curl;\Omega)$.
        \end{theorem}
        \item For a bounded Lipschitz domain $\Omega$ it holds that, if $\vec u$ in $H(\curl;\Omega)$ is such that
            $$ (\curl \vec u, \vec \phi) - ( \vec u, \curl \vec \phi) = 0$$
        for all $\vec \phi$ in $[C^\infty(\bar\Omega)]^3$, then $\vec u$ is actually in $H_0(\curl;\Omega)$.
        \item The trace operators here are two: 
            \begin{align*}
                \gamma_t \vec v &= \vec n \times \vec v, \\
                \gamma_T \vec v &= (\vec n \times \vec v) \times \vec n. 
            \end{align*}
        \item 
        \begin{theorem}
            For a bounded Lipschitz domain $\Omega$ it holds that the extension $\gamma_t: H(\curl;\Omega) \to H^{-1/2}(\partial\Omega)$ is bounded and linear, with the following integration by parts formula: 
                $$ (\curl \vec v, \vec\phi) - (\vec v, \curl\vec \phi) = \langle \gamma_t \vec v, \vec \phi\rangle \qquad\forall \vec v \in H(\curl;\Omega), \vec\phi \in \vec H^1(\Omega). $$
            **Note the bold space, which refers to a vector space. We will use this often. 
        \end{theorem}
        The operator $\gamma_t$ is not surjective simply because it is the limit of tangential vectors which will never have a normal component (at least intuitively). 
        \item Characterizing $\gamma_T$ is out of scope in this course. 
        \item 
        \begin{theorem}
            Consider a bounded Lipschitz domain $\Omega$. Then, 
                $$ H_0(\curl;\Omega) = \{ \vec v \in H(\curl;\Omega): \gamma_t \vec v = \vec 0\}. $$
        \end{theorem}
    \end{itemize}

\example{
One interesting context in which these trace operators show up is when considering normal or tangential boundary conditions. As we will see, one typically requires boundary information on all components of the solution on the boundary, but how this is done is highly context dependent. One nice example are \emph{slip} boundary conditions in fluid dynamics, where only the normal component of the fluid velocity is required to be 0: 
    $$ \vec u \cdot \vec n = 0.$$
This has to be complemented with conditions in the tangential direction. One possibility would be to prescribe some tangential velocity: 
    $$ (\ten I - \vec n \otimes \vec n)\vec u = (\vec n \times \vec u) \times \vec n = \vec v_\tau, $$
but one can also look at the tangential components of the stress tensor, thus generating a "tangential" Neumann boundary condition: 
    $$ (I - \vec n \otimes \vec n)[\sigma(\vec u)\vec n] = \vec g_\tau. $$
}

\subsection{Weak formulations}
A weak formulation refers to an integral form of a PDE, understood distributionally. This is typically a systematic procedure that should not be too difficult, and it helps in revealing what are the adequate boundary conditions for a given problem. The main tool for this will be the integration by parts formulas. Our test problem will be the Laplace problem, given by the $-\Delta$ operator. The minus sign will be better justified in the following section. Consider then the problem of finding $u$ such that 
    $$ -\Delta u = f \qquad \text{ in $\Omega$}. $$
Define an arbitrary smooth function $v$, then integration by parts yields
    $$ - \int_\Omega \Delta u v\,dx = -\int_{\partial\Omega}\gamma_D v \gamma_N \grad u \,ds + \int_\Omega \grad u \cdot \grad v\,dx $$
for all $v$. This function is typically called a \emph{test function}. The surface form suggest the boundary conditions: 
    $$ \int_{\partial\Omega}\underbrace{\gamma_D v}_\text{Dirichlet BC} \underbrace{\gamma_N \grad u}_\text{Neumann BC} \,ds,$$
so that we can have boundary conditions on the function itself  
    $$ u = g, $$
or on its normal derivative
    $$ \grad u \cdot \vec n = h. $$
This can be combined, so that for a given partition of the boundary into two sets $\Gamma_D$ and $\Gamma_N$ such that $\overline{\partial\Omega} = \overline{\Gamma_D}\cup\overline{\Gamma_N}$, one can have a Dirichlet boundary condition on $\Gamma_D$ and a Neumann boundary condition on $\Gamma_N$. For this type of boundary condition, one must define a solution space given by 
    $$ V_g = \{v \in H^1(\Omega): \gamma_D v = g \text{ on $\Gamma_D$}\}, $$
but let us focus first on spaces with null Dirichlet boundary condition ($V_0$). In this case, the boundary conditions will give
    $$ \int_{\partial\Omega}\gamma_Dv \gamma_N \grad u\,ds = \int_{\gamma_N} h \gamma_D v\,ds, $$
and thus the integral form of the equation will be given by
    $$ -\int_\Omega\Delta u v\,dx = -\int_{\Gamma_N}v h\,ds + \int_\Omega \grad u \cdot \grad v\,dx = \int_\Omega f v\,dx, $$
for all smooth $v$. Now, we note that (i) this formulation is well defined for $u,v$ in $H^1$ (and thus can be extended to hold for all $v$ in $H^1$ by density as long as $v$ satisfies the Dirichlet boundary conditions), (ii) the $\gamma_D$ operator has been omitted from the surface integral for convenience, and (iii) that the Dirichlet boundary condition does not appear anywhere in the formulation. This justifies naming Dirichlet boundary conditions \emph{essential}, and Neumann boundary conditions \emph{natural}. The \emph{weak formulation} of the problem thus refers to the following statement: Find $u$ in $V_0$ such that
    $$ (\grad u, \grad v)_{0,\Omega} = \langle f, v\rangle  + \langle h,  v\rangle \qquad \forall v\in V_0, $$
for given functions $f$ in $V_0'$ and $g$ in $(\gamma_D V_0)'$. Note the following: 
    \begin{itemize}
        \item The space of the solution and the test functions is the same. This is not mandatory, but it is common and can be better motivated by interpreting the Laplace problem as the first order equations related to the following minimization problem: 
            $$ \min_{v \in V_0} \int_\Omega |\grad v|^2\,dx . $$
        Then, one simply infers the spaces of each function from the definition of the Gateaux derivative. 
        \item The solution $u$ was formulated in a space without boundary condition. This is important because the regularity theory will depend on the solution space being a Hilbert space, and the space $V_g$ is not even a vector space as it is not closed under addition. This can be solved by defining adequate \emph{lifting} operators, i.e. a function $G$ in $H^1(\Omega)$ such that $\gamma_D G = g$ that allows us to write $u$ in $V_g$ as 
            $$ u = u_0 + G, $$
        where $u_0$ belongs to $V_0$. We can then rewrite the problem in $V_g$ as a problem in $V_0$ (and I encourage the reader to do this procedure at least once in their life). The existence of a lifting function in this case is given by the surjectivity of the Dirichlet trace, but it can be tricky in other contexts. This is also tricky in nonlinear problems, which justifies that nonlinear problems are typically studied with homogeneous boundary conditions.  
        \item We note that the Laplacian is now being interpreted as a \emph{distribution}, and thus the strong problem (including the boundary conditions) yields the definition of the \emph{action} of the distribution. In particular, this means that the action of the distribution naturally changes with the boundary conditions. This observation is fundamental to understand Discontinuous Galerkin methods, or other formulations defined on broken spaces (i.e. spaces that allow for discontinuities). 
    \end{itemize}

\example{
    We encourage the reader to try to comptue the weak formulation of the Poisson problem in mixed form. To do this, one must define the auxiliary variable $\vec \sigma \coloneqq \vec u$, so that the strong form of the problem now becomes
        $$
            \begin{aligned}
                -\dive \vec\sigma &= f &&\text{ in $\Omega$}, \\
                \vec \sigma - \grad u &= 0 &&\text{ in $\Omega$}, \\
                \gamma_D u &= g &&\text{ on $\Gamma_D$}, \\
                \gamma_N \vec\sigma &= h &&\text{ on $\Gamma_N$}. 
            \end{aligned}
        $$
    This problem will be studied in detail further ahead. 
}



\subsection{Poincarè inequalities and Lax-Milgram}


Using all of the previous definitions, we can finally look at actual problems and some first well-posedness results. For all of them, the Poincaré inequality will be fundamental.

\begin{lemma}[Generalized Poincaré inequality] There exists a positive constant $C$ such that for a non-empty portion of the subdomain $\Gamma \subseteq \partial\Omega$ it holds that
        $$ \| u \|_{0,\Omega} \leq C\left(| u |_{1,\Omega} + \left|\int_\Gamma u\,ds\right| \right) \qquad \forall u \in H^1(\Omega). $$
The result also holds in $L^p$ and $W_0^{1,p}$. 
\end{lemma}

We provide an incomplete proof to show some of the related techniques used to prove this result. Most developments come from \cite{brenner2008mathematical}.
\begin{proof}
    The main result to be used is the Bramble-Hilbert Lemma, which establishes (among other things) that if $B$ is a sufficiently big ball in $\Omega$ such that $\Omega$ is starred with respect to it, then the average over $B$ given by $\bar u = 1/|B|\int_B u\,dx$ satisfies
    $$ \| u - \bar u \|_{0,\Omega} \leq C| u |_{1,\Omega}, $$
where $C$ is a positive constant and $|\cdot|_{1,\Omega}$ is the $H^1$ semi-norm. The case $B=\Omega$ is known as the Friedrich's inequality. To recover Poincaré's inequality, one notes the two following properties: 
    $$ \|v \|_{0,\Omega} \leq \| v - \bar v\|_0 + \| \bar v\|_0 \leq C | v |_{1,\Omega} + \|\bar v\|_0,$$
where the first term was controlled using the Friedrich inequality. The second term is controlled by forcing another triangle inequality: 
    $$ \| \bar v\|^2_0 = \frac{|\Omega|}{|\Gamma|} \int_\Gamma \bar v\,ds \leq \frac{|\Omega|}{|\Gamma|}\left(\int_\Gamma (\bar v - v)\,ds + \left| \int_\Gamma v \,ds \right| \right)$$
and finally by using the trace inequality plus another application of the Friedrich inequality one gets the desired result. 
\end{proof}

The case in which $u$ is restricted to be in $H_0^1(\Omega)$ is the well-known Poincaré inequality. Note that in that case the boundary integral disappears. For the case of pure Neumann boundary conditions, we will require the following inequality.

\begin{lemma}[Poincaré-Wirtinger inequality]
    Consider an open connected bounded domain $\Omega$ as previously. Then, if we define the volumetric average as $\bar u = \int_\Omega u\,dx$, then it holds that
        $$ \| u - \bar u \|_{L^p(\Omega)} \leq C \| \grad u \|_{L^p(\Omega)}. $$
    \begin{proof}
        The proof has been taken from \cite{evans2022partial}. By contradiction, we assume that there exists a sequence of functions $(u_k)_k$ in $W^{1,p}(\Omega)$ such that
            $$ \| u_k - \bar u_k \|_p > k \| \grad u_k \|_p. $$
        We can renormalize the sequence by setting
            $$ v_k = \frac{u_k - \bar u_k}{\|u_k - \bar u_k\|_p}, $$
        so that $ \bar v_k = 0$ and $\| v_k \| = 1$. Our hypothesis yields
        $$ \| \grad v_k \|_p = \| \|u_k - \bar u_k\|_p^{-1} \grad (u_k - \bar u_k) \|_p < \frac 1 k,  $$
        which in particular means that $(v_k)_k$ is a bounded sequence in $W^{1,p}(\Omega)$. This means that $(v_k)_k$ has a weakly convergent sequence in $W^{1,p}(\Omega)$, and as this space is compactly embedded in $L^p(\Omega)$ (Rellich-Kondarchov), then there exists an element $v$ in $L^p(\Omega)$ such that, possibly up to a subsequence, $v_{k_j} \to v$. This implies that $\bar v=0$ and $\|v\|_p=1$. Using the bound on $\grad v_k$, one can also show that for smooth functions $\phi$, 
        $$ \int_\Omega v \partial_i \phi\,dx = \lim_k \int v_k \partial_i \phi\,dx = - \lim_k \int_\Omega \partial_i v_k \phi = 0. $$
        This establishes that $\grad v=0$, which implies that $v$ is constant (it is not trivial to check that null weak derivatives implies being a constant), and the null average condition yields that $v=0$. This contradicts the initial hypothesis. 
    \end{proof}
\end{lemma}

Probably the most important consequence of this is that the semi-norm given by $| x | \coloneqq \|\grad x\|_{0,\Omega}$, sometimes referred to as the $H_0^1$ semi-norm, is equivalent to the $H^1$ norm in the following cases: (i) homogeneous Dirichlet boundary conditions, (ii) homoegeneous Neumann boundary conditions, and (iii) mixed homogeneous Dirichlet and Neumann boundary conditions. Verifying this is a simple but fundamental excercise. A fundamental property to be verified is the following: a bilinear form $a(\cdot, \cdot)$ defined on a Hilbert space $X$ is said to be elliptic if there exists a constant $\alpha$ such that
        $$ a(x, x) \geq \alpha \| x \|^2_X \qquad \forall x\in X. $$

\begin{lemma}[Lax-Milgram] Consider a bounded bilinear form $a: H\times H\to \R$ defined on a Hilbert space $H$ that is elliptic with constants $C$ and $\alpha$ respectively, and a linear functional $f$ in $H'$. Then, there exists a unique $u$ in $H$ such that 
    $$ a(u, v) = f(v) \qquad \forall v \in H. $$
This solution is continuous with respect to the data, in the sense that there exists a positive constant $C$ such that 
    $$ \| u\|_H \leq \frac C \alpha \| f \|_{H'} .$$
\end{lemma}



\paragraph{The Poisson problem} Consider $f$ in $H^{-1}(\Omega)$ and $g$ in $H^{1/2}(\Gamma)$ with $\Gamma\coloneqq \partial\Omega$. The Poisson problem in strong form is given as the following PDE: 
    \begin{align*}
        -\Delta u  &= f \qquad \tin\quad\Omega\\
        \gamma_0 u &= g \qquad \ton\quad \Gamma.
    \end{align*}
Note that the strong form must be understood in the distributional sense, i.e. as an equation in $H^{-1}(\Omega)$. To derive the weak formulation, consider a function $v$ in $H_0^1(\Omega)$, then using the boundary conditions we obtain that 
    $$ -\langle \Delta u,v\rangle = (\grad u, \grad v),$$
where $(\cdot, \cdot)$ is the $L^2(\Omega)$ product. Thus the weak formulation reads: Find $u$ in $H_0^1(\Omega)$ such that 
    $$ \int_\Omega \grad u\cdot \grad v\,dx = \langle f, v\rangle \qquad \forall v\in H_0^1(\Omega).$$
This problem can be shown to be well-posed using Lax-Milgram's lemma and the Poincarè inequality. Small excercise: Extend the proof to the case of non-homogeneous Dirichlet boundary conditions.

In the case of having a boundary condition defined only on a portion $\Gamma_D$ of the boundary, the formulation changes, because (i) we need further information regarding the Neumann trace on the complement of the boundary, (ii) the test space looks different. In particular, we define the solution space given by 
    $$ V_0 = \{v\in H^1(\Omega): \quad v = 0 \quad\text{ on \Gamma_D}\}, $$
which using the generalized Poincaré can be shown to still satisfy an ellipticity estimate. 

\paragraph{The pure Neumann problem} In general, having Neumann boundary conditions is problematic for two reasons: It results in a \emph{data compatibility} condition and (ii) it results in having a non-trivial kernel in the problem. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Finite elements for elliptic problems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To be fair, we will actually consider a Galerkin scheme, which can be described as follows: Consider the problem of finding $u$ in $H$ such that
    $$ [...] $$


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Beyond ellipticity}\label{section:beyond-ellipticity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Inf-sup conditions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Saddle point problems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discretization of saddle point problems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Beyond linearity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fixed point theorems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Monotone operators}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Time dependent problems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Faedo-Galerkin and the method of lines}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Space and time discretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Poroelasticity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Equilibrium equations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constitutive modeling}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Darcy and Biot equations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{main}
\bibliographystyle{alpha}
\end{document}

