The theory of well-posedness for linear problems, centered on the Lax-Milgram lemma~\ref{lemma:lax-milgram}, provides a robust framework for proving existence and uniqueness of solutions. Its power lies in its direct application to any problem whose weak formulation can be expressed through a continuous and elliptic bilinear form $a(u,v)$. Many physical models, however, are inherently nonlinear, leading to equations where this linear structure is lost. Consider, for instance, a semilinear problem of the form
\begin{equation}\label{eq:semilinear-example}
    \begin{aligned}
        -\Delta u + F(u) &= f &&\tin \Omega,  \\
        u &= 0 && \ton \partial\Omega.
    \end{aligned}
\end{equation}
Its weak formulation, $(\nabla u, \nabla v) + (F(u),v) = (f,v)$, results in a form $a(u;v)$ that depends nonlinearly on $u$, and consequently, the Lax-Milgram lemma is not applicable. Thus, we need a new set of tools to establish the well-posedness of such problems.

This chapter introduces the foundational methods of nonlinear functional analysis. Unlike in the linear setting, there is no single, universally applicable theorem for existence and uniqueness. Instead, we will build a toolkit of three distinct and powerful approaches, each suited for a different class of nonlinear problems.

We will begin with a \emph{local analysis} based on the inverse function theorem, which provides conditions for a unique solution to exist in a neighborhood of a known solution. We then move to more global methods founded on the \emph{fixed point theorems} of Banach, Brouwer, and Schauder, which provide existence results under assumptions of contractivity or compactness. Finally, we will explore the theory of \emph{monotone operators}, a nonlinear generalization of positive definite operators, and establish the Minty-Browder theorem, a cornerstone result that guarantees the existence of solutions for coercive and monotone operators.

\section{Local analysis}\label{sec:local-analysis}
Consider the set
\[ \inv(X,Y) = \left\{ A\in \mathcal{L}(X,Y): A \text{ is invertible} \right\}. \]

Then, we have the following local inversion result
\begin{theorem}[Local invertibility]\label{thm:local-invertibility}
    Consider $F\in C^1(X,Y)$, $F'(u^*)\in\inv(X,Y)$. Then, $F$ is locally invertible at $u^*$ with $F^{-1}\in C^1$. In fact,
    \begin{equation}
        dF^{-1}(v) = (F'(u))^{-1}, \quad u=F^{-1}(v).
    \end{equation}
\end{theorem}

\example{
    Consider the nonlinear problem 
    \begin{equation*}
        \begin{aligned}
            \Delta u + u^p  &= h, &&\tin \Omega,\\
            u&=0, &&\ton\partial\Omega,
        \end{aligned}
    \end{equation*}
    where $p>1$, $h\in H^{-1}(\Omega)$, so $F:X\to Y$ with $X\coloneqq H_0^1(\Omega)$ and $Y\coloneqq H^{-1}(\Omega)$. We compute
    \begin{equation}
        dF(u)[w] = -\Delta w + p u^{p-1} w.
    \end{equation}
    We want to show that $dF(u)\in\inv(X,Y)$. That is, for a given $g\in Y$, we need to prove that there exists a $w\in X$ that solves
    \begin{equation*}
        a(w,v) = (\nabla w, \nabla v) + p(u^{p-1}w,v) = \langle g,v \rangle \quad \forall v\in X.
    \end{equation*}
    We have proven many times before the bound for the first term, so let us focus in the second:
    \begin{equation*}
        |(u^{p-1}w,v)|\leq \|u^{p-1} \|_{L^\infty} \|w\|_{0} \|v \|_{0}.
    \end{equation*}
    Note that if $u\geq 0$, then $a$ is elliptic. Then, if $u\in X\cap L^\infty \cap \{u\geq 0\}$ we have that $dF(u)\in\inv(X,Y)$ and we can use Theorem~\ref{thm:local-invertibility}.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fixed point theorems}\label{sec:fixed-point-theorems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The main idea is that certain nonlinearities would be less terrible if we could fix one of the functions. For example, consider the problem of finding a function $u$ such that
\begin{equation}
    \begin{aligned}
        -\Delta u &= \sin (u) &&\tin \Omega,  \\
        u &= 0 && \ton  \partial\Omega.
    \end{aligned}
\end{equation}
This problem is non-linear, but given $w$, finding $u$ such that
\begin{equation}\label{eq:fixed-point-sin}
     -\Delta u = \sin(w)
\end{equation}
is linear on $u$. It also induces a mapping $w \mapsto T(w) = u$ such that, \emph{if it has a fixed point} $u=T(u)$, this fixed point solves the initial problem. For this section, we follow~\cite{ciarlet2013linear,pata2019fixed}.
\begin{definition}[Non-expanding and contractive maps]\label{def:nonexpanding-contractive-maps}
    A Lipschitz function $f:X\to X$
    \begin{equation*}
        ||f(x)-f(y)||\leq L||x-y||
    \end{equation*}
    is said to be \emph{non-expansive} if $L=1$, and a \emph{contraction} if $L<1$.
\end{definition}
\begin{theorem}[Banach fixed point]\label{thm:banach-fixed-point}
    Let $X$ be a complete metric space and $f$ a contraction of constant $\lambda<1$. Then $f$ has a unique fixed point.
    \begin{proof} 
    Set $x^{n+1}=f(x^n)$, which we typically refer to as a \emph{Picard iteration}, and consider some initial $x^0\in X$. By induction we get that
    \begin{tightalign*}
        \|x^{n+1}-x^n\| &\leq \lambda \|x^n-x^{n-1}\|\leq \underbrace{\dots}_\text{induction} \lambda^n\|x^1-x^0\|\\
        \Rightarrow  \| x^{n+m}-x^n\| &\leq \|x^{n+m}-x^{n+m-1}\|+ \cdots = \|x^{n+1}-x^n\|\\
        &\leq (\lambda ^{n+m-1}+\cdots+\lambda^n)\|x^1-x^0\|\\
        &=\lambda^n\left(\displaystyle\sum_{j=0}^m\lambda^j\right)\|x^1-x^0\|\\
        &\leq\lambda^n\left(\displaystyle\sum_{j=0}^\infty\lambda^j\right)\|x^1-x^0\|\\
        &=\lambda^n\dfrac{1}{1-\lambda}\|x^1-x^0\|
    \end{tightalign*}
    which implies that $\{x_n\}_n$ is Cauchy, and since $X$ is complete, this gives $x_n\to \bar{x}$. Finally, since $f$ is continuous, $f(\bar{x}) = \lim_{n\to \infty}f(x_n) = \lim_{x\to\infty}x_{n+1}=\bar{x}$.
    \end{proof}
\end{theorem}
We note that if $X$ is compact, then this theorem can be extended to \emph{weak} contractions, i.e. $\|f(x)-f(y)\|< \|x-y\|$.\\

\example{
    Consider problem~\eqref{eq:fixed-point-sin}, where we note that the Lax-Milgram lemma~\eqref{lemma:lax-milgram} immediately yields the existence of a unique solution $u$ for each $w$. From the mean value theorem~\eqref{thm:mean-value-banach} one has that differentiable functions satisfy $f(x) - f(y) = f'(\xi)(x-y)$ for some $\xi$ in $(x,y)$. This in particular shows that $\sin$ is a Lipschitz function as $\sup_{\xi\in \R}\sin(\xi) = 1$. Using this fact we can compute the following for some given $w_1, w_2$:  
    \begin{equation*}
        -\Delta(T(w_1)-T(w_2))= \sin(w_1)-\sin(w_2),
    \end{equation*} 
    and the a priori estimate gives
    \begin{equation*}
        \|T(w_1)-T(w_2)\|\leq C\|\sin(w_1)-\sin(w_2)\|\leq C \|w_1-w_2\|.
    \end{equation*}
    Then, if $C<1$ there exists $\bar u$ such that $T(\bar u)= \bar u$. 
}
\begin{theorem}[Brouwer fixed point]\label{thm:brouwer-fixed-point}
    Set $K$ a non-empty, compact and convex subset of a finite-dimensional Banach space. Then, every continuous function $f:K\to K$ has a fixed point.
\end{theorem}
\example{
    Let's go back to~\eqref{eq:fixed-point-sin}, and consider its Galerkin scheme for some given $w_h$:
    \begin{equation*}
        (\grad u_h,\grad v_h)=(\sin(w_h),v_h) \qquad \forall v_h \in V_h
    \end{equation*}
    where $\dim(V_h)<\infty$. Then, from the a priori bound we have 
    \begin{equation*}
        \|u_h\|_{1}\leq C||\sin (w_h)||< C
    \end{equation*}
    using that $|\sin (x)|\leq 1$. This implies that $u_h$ is contained in a finite-dimensional ball of $V_h$, $\mathcal{B}(0,r), r=C$. To show that it is continuous, we consider two functions $w_1,w_2$ in $V_h$, and obtain the difference equation setting $u_i=T(w_i)$: 
    \begin{equation*}
        (\grad [u_1 - u_2], \grad v_h) = (\sin(w_1) - \sin(w_2), v_h) \qquad \forall v_h \in V_h.
    \end{equation*}
    Again, using the a priori bound we use $\sin'(x) = \cos(x) \leq 1$ to obtain 
    \begin{equation*}
        \| u_1 - u_2\|_1 \leq \|w_1 - w_2\|_0.
    \end{equation*}
    This yields that $T_h:\mathcal{B}(0,r)\to \mathcal{B}(0,r)$ is continuous and thus has at least one fixed point.
}

From the previous result we can extend to $h\to 0$ through compactness. This generalization is known as Schauder's fixed point theorem, which has two forms.
\begin{theorem}[Schauder fixed point]\label{thm:schauder-fixed-point}
    \begin{enumerate}
        \item Set $K$ a compact, convex subset of $X$ normed space, and $f:K\to K$ a continuous mapping. Then $f$ has at least one fixed point.
        \item Let $\mathcal{C}$ a closed, convex subset of a Banach space $X$, and $f:\mathcal{C}\to \mathcal{C}$ continuous s.t. $\overline{f(\mathcal{C})}$ is compact. Then $f$ has at least one fixed point.
    \end{enumerate}
\end{theorem}
\example{ 
    Consider again~\eqref{eq:fixed-point-sin}, with its induced Picard mapping $w \mapsto u\coloneqq T(w)$ given by 
    \begin{equation}
        \begin{aligned}
            -\Delta u &= \sin(w) && \tin \Omega,\\
            u&=0 &&\ton \partial\Omega,
        \end{aligned}
    \end{equation}
    whose solution is guaranteed by Lax-Milgram's lemma, with $u$ in $H_0^1(\Omega)$. From the a priori bound $\|u||\leq\dfrac{1}{\alpha}\|f\|$ we get $\|u\|_1\leq C\|\sin(w)\|_0\leq C|\Omega|=:r_0$. Our candidate set will be $\mathcal C = \bar{B}(0,r_0)\subset L^2(\Omega)$ which is closed, convex and contained in both $L^2(\Omega)$ and $H_0^1(\Omega)$. To prove that $\overline{T(\mathcal C)}$ is compact, consider a sequence $(w_k)_{k>1}$ in $\mathcal C$, which yields $(v_k)_{k>1}=(T(w_k))_{k>1}$ in $\mathcal C$ and additionally $v_k\in H_0^1(\Omega)$ for all $k$. Now, we note that we have a sequence in $H^1$, and as the unit ball is weakly compact, we obtain that is has some weakly convergent sequence in $H^1$. The final detail is that, if we consider $\mathcal C$ to be a ball in $L^2$, then the mapping $T$ can be written additionally as $T\circ i$, where $i:L^2(\Omega)\to H^1(\Omega)$ is the compact embedding of $H^1$ into $L^2$. We thus have that $T\circ i$ is the composition of a continuous and a compact mapping, and thus a compact map. This in particular implies that our sequence $(v_k)$ has a strongly convergent subsequence in $L^2$. This concludes that $T:L^2(\Omega)\to L^2(\Omega)$ is compact, and thus we can use Schauder's fixed point theorem to show the existence of a fixed point $\bar u = T(\bar u)$ in $L^2(\Omega)$. Naturally, the equation itself reveals that actually $\bar u$ belongs to $H_0^1(\Omega)$, so we have some additional regularity on the fixed point we just found.
}
A consequence of the Schauder fixed point theorem is the Schaefer fixed point theorem. We state it for completeness without proof.
\begin{theorem}[Schaefer fixed point]\label{thm:schaefer-fixed-point}
    Let $X$ be a Banach space and $f:X\to X$ a compact operator. If 
    \begin{equation}
        \{x\in X: \sigma f(x)=x,\sigma\in[0,1]\}\subset B(0,r)
    \end{equation}
    for some $r>0$, then there exists a fixed point of $f$.
\end{theorem}
    
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Monotone operators}\label{sec:monotone}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Among nonlinear operators, there is a class of operators that provide efficient means to prove existence and uniqueness of solutions. These operators are called \emph{monotone}, and they have been extensively studied. We follow this section closely from~\cite{ciarlet2013linear}.
\begin{definition}[Monotone operator]\label{def:monotone-operator}
    Let $V$ be a normed vector space and $\langle\cdot,\cdot\rangle$ its duality pairing. An operator $A:V\to V'$ is called \emph{monotone} if
    \begin{equation}
        \langle A(v)-A(u),v-u\rangle \geq 0 \qquad \forall u,v\in V,
    \end{equation}
    and \emph{strictly monotone} if the inequality is strict for $u\neq v$.
\end{definition}
We note that if $f:X\to \R$ is a convex and differentiable function, then $\nabla f$ is a monotone operator. Indeed, from convexity applied to $x,y\in X$ and adding we get
\begin{tightalign*}
    f(x) &\geq f(y) + \langle \nabla f(y), x-y\rangle\\
    f(y) &\geq f(x) + \langle \nabla f(x), y-x\rangle\\
    \implies 0 &\geq \langle \nabla f(y)-\nabla f(x),x-y\rangle \\
    \implies 0&\leq \langle \nabla f(x)-\nabla f(y),x-y\rangle.
\end{tightalign*}
We can also build monotone operators from an elliptic operator. If $A:V\to V'$ is an operator such that $a(u,v) = \langle A(u),v\rangle$ is continuous (linear) and elliptic, then $A$ is strictly monotone:
\begin{tightalign*}
    \langle A(u)-A(v),u-v\rangle &= \langle A(u-v), u-v\rangle\\
    &= a(u-v,u-v)\\
    &\geq \alpha \|u-v\|^2 \geq 0,
\end{tightalign*}
and the inequality is strict for $u\neq v$ since $\|\cdot\|$ is a norm. 

A very useful property can be derived from the monotonicity in the case that $V$ is complete.
\begin{theorem}\label{thm:lipschitz-property-monotone}
    If $V$ is a real Banach space and $A:V\to V'$ is monotone, then $A$ is locally bounded, i.e. for any $u\in V$ there exists $r=r(u)>0$ and $\rho=\rho(u)>0$ such that
    \begin{equation*}
        \|u-v\|\leq r \implies \|A(u)-A(v)\|\leq \rho,
    \end{equation*}
    which similar to a Lipschitz property. Moreover, if $A$ is linear, then $A$ is continuous.
\end{theorem}

We now introduce the notion of coercive ($\neq$ elliptic) and hemicontinuous operators.
\begin{definition}[Coercive operator]\label{def:coercive-operator}
    An operator $A:V\to V'$ is \emph{coercive} if 
    \begin{equation*}
        \lim_{\|v\|\to \infty} \frac{\langle A(v),v\rangle}{\|v\|} = +\infty.
    \end{equation*}
\end{definition}
\begin{definition}[Hemicontinuous operator]\label{def:hemicontinuous-operator}
    An operator $A:V\to V'$ is \emph{hemicontinuous} if for every $u,v,w\in V$ there exists $t_0=t_0(u,v,w)>0$ such that the map 
    \begin{tightalign*}
        \varphi:(-t_0,t_0)&\to \R\\
        t&\mapsto \langle A(u+tv),w\rangle
    \end{tightalign*}
    at $t=0$. 
\end{definition}
Hemicontinuity is a weaker property than continuity, as we shall prove in the following theorem.
\begin{theorem}
    If $A:V\to V'$ is continuous, then it is hemicontinuous.
    \begin{proof}
        Let $\varphi(t)=\langle A(u+tv),w\rangle$. Then, we get
        \begin{tightalign*}
            |\varphi(t)-\varphi(0)| &= |\langle A(u+tv),w\rangle - \langle A(u),w\rangle|\\
            &\leq |\langle A(u+tv)-A(u),w\rangle| \tag{Linearity of $A$}\\
            &\leq \|A(u+tv)-A(u)\|_{V'} \|w\|_V. \tag{Cauchy-Schwarz}
        \end{tightalign*}
        Thus, as $\|(u+tv)-u\|_V=\|tv\|\leq |t|\|u\|_V$, then $u+tv\overset{t\to 0}{\to} u$ in norm, and since $A$ is continuous, we get $A(u+tv)\overset{t\to 0}{\to} A(u)$ in $V'$. Thus, taking limit as $t\to 0$ we immediately conclude that $\varphi$ is continuous at $t=0$.
    \end{proof}
\end{theorem}
Now that we have defined hemicontinuous operators, we are ready to introduce a theorem that gives sufficient conditions that guarantee the surjectivity of a hemicontinuous monotone operator.
\begin{theorem}[Minty-Browder]\label{thm:minty-browder}
    Let $V$ a real, separable and reflexive Banach space (e.g. $L^p$ for $p>1$), and $A:V\to V'$ a coercive and hemicontinuous monotone operator. Then, $A$ is surjective, i.e. given any $f\in V'$ there exists $u\in V$ such that $A(u)=f$. Moreover, if $A$ is strictly monotone, then $A$ is also injective, and thus there exists a unique solution for $A(u)=f$.
\end{theorem}
A common example of an operator that can be analyzed via the Minty-Browder theorem is the \emph{p-Laplacian} $-\Delta_p$, which for $p\geq 1$ is defined as the map
\begin{tightalign*}
    -\Delta_p : W_0^{1,p}(\Omega)&\to W^{-1,q}(\Omega) = (W_0^{1,p}(\Omega))'\\
    v&\mapsto -\Delta_p v = -\nabla\cdot(|\nabla v|^{p-2}\nabla v),
\end{tightalign*}
where $q$ is the conjugate exponent of $p$, such that $1/p + 1/q = 1$. Note that for arbitrary $u,v\in W_0^{1,p}(\Omega)$ the duality is given by 
\begin{equation*}
    \langle \Delta_p u, v\rangle = \langle \nabla\cdot (|\nabla u|^{p-2}\nabla u), v\rangle = -\langle |\nabla u|^{p-2}\nabla u, \nabla v\rangle = -\int_\Omega |\nabla u|^{p-2}\nabla u\cdot\nabla v dx,
\end{equation*}
which is well-defined by Hölder's inequality:
\begin{equation*}
    \left|\int_\Omega |\nabla u|^{p-2}\nabla u\cdot\nabla v dx\right| \leq \|\nabla u\|^{p-1}_{W_0^{1,p}(\Omega)}\|\nabla v\|_{W_0^{1,p}(\Omega)},
\end{equation*}
and noting that $w\mapsto \|\nabla w\|_{W_0^{1,p}(\Omega)}$ is a norm on $W_0^{1,p}(\Omega)$. We define now the functional
\begin{tightalign*}
    \Psi:W_0^{1,p}(\Omega)&\to \R\\
    u&\mapsto \Psi(u) \coloneqq  \frac{1}{p}\int_\Omega |\nabla u|^p dx.
\end{tightalign*}
This functional is Gâteaux-differentiable, and we explicitly compute its the Gâteaux derivative for $u,v\in W_0^{1,p}(\Omega)$:
\begin{tightalign*}
    d\Psi(u)[v] &= \left.\frac{d}{d\varepsilon}\right|_{\varepsilon=0} \Psi(u+\varepsilon v)\\
    &= \left.\frac{d}{d\varepsilon}\right|_{\varepsilon=0} \frac{1}{p}\int_\Omega \underbrace{|\nabla (u+\varepsilon v)|^p}_{|\nabla(u+\varepsilon v)\cdot \nabla(u+\varepsilon v)|^{p/2}} dx\\
    &= \left.\frac{1}{p}\int_\Omega \frac{p}{2} |\nabla (u+\varepsilon v)\cdot\nabla(u+\varepsilon v)|^{p/2-1}\cdot 2(\nabla u\cdot\nabla v) dx\right|_{\varepsilon = 0}\\
    &= \int_\Omega |\nabla u\cdot\nabla u|^{\frac{p-2}{2}}\nabla u\cdot\nabla v dx\\
    &= \int_\Omega |\nabla u|^{p-2}(\nabla u\cdot\nabla v) dx\\
    &= \langle -\Delta_p u, v\rangle,
\end{tightalign*}
and thus $d\Psi = -\Delta_p$. This operator is hemicontinuous: take $t\in \R$ and $u,v,w\in W_0^{1,p}(\Omega)$, and define $\varphi(t) = \langle A(u+tv),w\rangle$. With this, we have
\begin{tightalign*}
    |\varphi(t)-&\varphi(0)| = \left|\int_\Omega \left(|\nabla (u+tv)|^{p-2}\nabla (u+tv) - |\nabla u|^{p-2}\nabla u\right)\cdot \nabla w dx \right|\\
    &= \left|\int_\Omega \left((|\nabla u + t\nabla v|^{p-2} - |\nabla u|^{p-2})\nabla u \cdot \nabla w + t|\nabla u + t\nabla v|^{p-2} \nabla v\cdot \nabla w \right) dx\right|\\
    &\leq \int_\Omega \left(|\nabla u + t\nabla v|^{p-2} - |\nabla u|^{p-2}\right)|\nabla u||\nabla w| dx + |t|\int_\Omega |\nabla u + t\nabla v|^{p-2}|\nabla v||\nabla w| dx,
\end{tightalign*}
and taking the limit $t\to 0$ we conclude that $-\Delta_p$ is hemicontinuous. Moreover, since $\Psi$ is strictly convex, we obtain that for all $u\neq v \in W_0^{1,p}(\Omega)$, it holds that
\begin{equation*}
    \langle \Delta_p u - \Delta_p v, u-v\rangle < 0 \implies \langle -\Delta_p u - (-\Delta_p v), u-v\rangle > 0,
\end{equation*}
which implies that $-\Delta_p$ is strictly monotone. To prove coercivity, we take a nonzero $u\in W_0^{1,p}(\Omega)$ and get
\begin{tightalign*}
    \frac{\langle A(u),u\rangle}{\|u\|_{W_0^{1,p}(\Omega)}} &= \frac{1}{\|u\|_{W_0^{1,p}(\Omega)}}\int_\Omega \underbrace{|\nabla u|^{p-2}(\nabla u\cdot\nabla u)}_{|\nabla u|^p} dx\\
    &= \frac{1}{\|u\|_{W_0^{1,p}(\Omega)}}\|u\|_{W_0^{1,p}(\Omega)}^p\\
    &= \|u\|_{W_0^{1,p}(\Omega)}^{p-1}\xlongrightarrow{\|u\|_{W_0^{1,p}(\Omega)}\to\infty}\infty,
\end{tightalign*}
since $p>1$, which proves that $-\Delta_p$ is coercive. Since $W_0^{1,p}(\Omega)$ is separable and reflexive for $p>1$, the conditions of the Minty-Browder theorem~\ref{thm:minty-browder} are satisfied, and thus we conclude that $-\Delta_p$ is bijective, i.e. for any $f\in W^{-1,q}(\Omega)$ there exists a unique $u\in W_0^{1,p}(\Omega)$ such that $-\Delta_p u = f$.
